Implementing Reactive Systems Directly\\with I/O Automata in ioa++

### 2012-02-23 papers/paper2/case_study.tex
Removing material.

In this section, we illustrate how ioa++ can be used to build real distributed systems.
In practice, distributed systems are designed using a layered approach where the highest layers contain (often semantically rich) application logic.
The lower layers provide services such as TCP/UDP sockets (for communicating between end systems) and timers (for detecting time-outs and performing periodic tasks), e.g., \cite{schmidt2001c++},~\cite{schmidt2002c++}.
The intermediate layers, i.e., frameworks like ioa++, contain generic coordination mechanisms that serve to abstract away the low-level details from the application logic.

>>>>

In this section, we illustrate how ioa++ can be used to build real distributed systems.
In practice, distributed systems are designed using a layered approach where the highest layers contain (often semantically rich) application logic.
The lower layers provide services such as TCP/UDP sockets (for communicating between end systems) and timers (for detecting time-outs and performing periodic tasks), e.g., \cite{schmidt2001c++},~\cite{schmidt2002c++}.
The intermediate layers contain generic coordination mechanisms that serve to abstract away the low-level details from the application logic.

### 2012-02-23 papers/paper2/component_model.tex
Removing material.

The Input/Output (I/O) automata model was developed by Lynch and Tuttle to model reactive concurrent and asynchronous systems~\cite{lynch1987hierarchical},\cite{lynch1996distributed}.
An I/O automaton consists of state variables and a set of atomic actions that manipulate the state variables.
An action is only permitted to manipulate the state variables of the automaton to which it belongs.
An action is either an internal action, an output action, or an input action.
An internal action only changes the state of the automaton.
An output action changes the state of the automaton and produces a signal or value that may be consumed by one or more input actions.
An input action changes the state of the automaton when it receives a signal or value produced by an output action.
\ifjournal
The set of actions associated with an automaton is known as the automaton's \emph{signature}~\cite{lynch1996distributed}.
\fi
An automaton's output and internal actions constitute its \emph{local signature}~\cite{lynch1996distributed}.
\ifjournal
Often, a local action consists of a precondition and an effect.
The precondition is a predicate over the state variables that enables the effect when selected for execution.
The effect is a function that computes the next state of the automaton from the current state.
An input action presented in this style consists solely of an effect.
\fi
An automaton's output and input actions constitute its \emph{external signature}~\cite{lynch1996distributed}.

An automaton communicates with other automata through the actions in its external signature.
I/O automata are \emph{composed} by (1)~concatenating the state variables of the constituent automata, (2)~consolidating the effects of all input actions with the same name, and (3)~incorporating the effects of input actions into output actions with the same name.
Composing automata that have identically named output actions would be ambiguous and thus is not allowed.
The result of composition is a new automaton, so that the principles for reasoning about the behavior of a composed system are the same as the principles for reasoning about a single automaton.
Often, the properties of the resulting aggregate automaton can be reasoned about using only the properties of the constituent automata.
%% A useful operation in composite automata is \emph{hiding}~\cite{lynch1996distributed} which converts an output action to an internal action.

Execution in the I/O automata model consists of non-deterministically and repeatedly selecting a local action and then if the precondition is true applying the effect.
The scheduler is assumed to be \emph{fair,} meaning that a local action is guaranteed to be selected (but not necessarily executed) infinitely often.
Executing internal actions and output actions that are not composed with any input actions is straightforward: in one atomic step, the precondition is evaluated and the effect is applied if and only if the precondition is true.
Executing a composed output action consists of atomically evaluating the precondition, and if it is true applying the effect of the output action and all composed input actions using the value produced by the output action.
The atomic relationship between composed output actions and input actions means that multiple automata may change state atomically when executing an output action.
The guarantee of atomicity allows a developer to relate the state of one automaton to another, resulting in well-defined interactions that facilitate composition.

@@

\begin{figure*}
\center
\includegraphics[width=\textwidth]{system_model}
\caption{Example event detection system.}
\label{sys_model}
\end{figure*}
Component frameworks that support \emph{recursive encapsulation} allow components to contain other components.
Recursive encapsulation is at the heart of both bottom-up design strategies where smaller components are composed to produce larger ones, and top-down design strategies in which larger components are decomposed into smaller ones.
The contained component is the \emph{child} while the containing component is the \emph{parent}.
The resulting hierarchy helps developers organize programs by separating concerns and facilitates the reuse of existing components.
A distinction is made between the \emph{name} given to an instance and the \emph{type} of a child component to allow a parent component to contain multiple children of the same type.

\paragraph*{Motivation for I/O automata based components}
%The I/O automata formalism was developed to model reactive systems~\cite{lynch1987hierarchical} and has been used in the design and verification of concurrent and distributed systems.
I/O automata are a good foundation for such components because they have independent state, straightforward interfaces, and precise semantics for event generation, distribution, and consumption resulting in well-defined interactions under composition.
\emph{Hierarchical decomposition}, the modeling dual of recursive encapsulation, is an established technique for I/O automata~\cite{lynch1987hierarchical},~\cite{lynch1994atomic}.
For convenience, we extend the notion that a parent automaton is the composition of its children to include an anonymous child that contains state and actions attributed to the parent.
The \emph{root automaton} is the automaton at the top of the hierarchy and represents the entire system as composition is applied recursively up the hierarchy.

@@

\paragraph*{Explicit composition}
Composition in the I/O automata model is based on actions having the same name and can be applied to an arbitrary number of automata.
However, when using recursive encapsulation, developers must have the ability to compose component automata regardless of the naming convention used by the original developers.
To prepare for dynamic composition, we limit the scope of composition to associating a single output action with a single input action.
Such an association is called a \emph{binding} and consists of a pair $(output, input)$ where $output$ is the name of the output action and $input$ is the name of the input action.

%% This is only true for static systems.  Dynamic system can do crazy things.
Each component automaton contains a set of bindings referring to its own actions, the actions of its children, the actions of its children's children, etc.
The automaton that prescribes a binding is said to \emph{own} that binding.
The global set of bindings can be used to rename input actions before applying composition as defined by the I/O automata model.
A map for renaming inputs is generated by replacing action names with their fully-qualified names in all bindings by tracing from the root automaton.
For example, the fully-qualified name of the $in?(s)$ action of Converter C in Figure~\ref{sys_model} is $[root].C.in?(s)$.
Composition as defined by the I/O automata model can be applied after using the map to rename all input actions to their corresponding output action.
The map of bindings for the event detection system shown in Figure~\ref{sys_model} is: 
\begin{displaymath}
\begin{split}
\{ ([root].out!(s), [root].C.in?(s)),\\
   ([root].C.out!(t), [root].D.in?(t)),\\
   ([root].D.detect!, [root].detected?) \}.
\end{split}
\end{displaymath}

\paragraph*{Binding rules}
To support name-based composition under the I/O automata model, the set of bindings must adhere to four essential rules.
First, the output action and input action of a binding must agree on the type being produced and consumed.
For example, an input action consuming an integer value can only be bound to an output action that produces an integer value.
The system shown in Figure~\ref{sys_model} follows this rule since $(out!(s), in?(s))$ agree on type $s$, $(out!(t), in?(t))$ agree on type $t$, and $(detect!, detected?)$ agree on the absence of a value (i.e., use of a signal).
Second, an input action can be bound to at most one output action.
Composing an input action with multiple output actions would introduce ambiguity into the relationship between the states of the automata and is therefore prohibited by the model.
A binding map where all bindings that mention the same input action are equivalent satisfies this rule.
For dynamics, which are discussed in Section~\ref{dynamics}, we must strengthen this rule to say that an input action can appear at most once in the global binding map.
This stronger rule means that an input is the sink of at most one binding and that every binding has exactly one owner.
Third, an output action cannot be bound to more than one input action in the same automaton:  the state of the automaton containing the input actions would not be well-defined after the output is executed since the input action effects would be applied in an undefined order.
%% Let $\pi$ be a prefix operator that strips the action name (leaving the path and hence the automaton) of a fully-qualified action name.
%% The following must be true for the global binding map $B$:  $\langle \forall o : (o, z) \in B :: \langle \forall i,j : (o, i) \in B \land (o, j) \in B \land i \neq j:: \pi (i) \neq \pi (j) \rangle \rangle$.
Fourth, an output action cannot be bound to an input action in the same automaton.
Allowing such bindings would result in undefined behavior since the order of the output and input effects is undefined.
The I/O automata model prevents this by requiring that all actions belonging to the same automaton have a unique name.
To elaborate, we usually think about the output action occurring before the input action because a value must be generated.
However, the I/O automata model admits an interpretation where the value is generated first and the effects of the output action and input action are applied non-deterministically.
Graphically, an arrow originating in one component automaton must terminate in another component automaton as Figure~\ref{sys_model} illustrates.
%% To adhere to this rule, the following must be true for the global binding map $B$:  $\langle \forall o, i : (o, i) \in B :: \pi (o) \neq \pi (i) \rangle$.

\paragraph*{Concurrent execution}
Recall that composing I/O automata consists of concatenating state variables and folding input actions into output actions and that the result of composition is a single equivalent automaton.
Since the state of each automaton is independent, opportunities exist for concurrent execution if the state of each automaton is preserved and a composition is not reduced to a single equivalent automaton.
Each local action implies a set of automata which in turn implies a set of state variables that might be modified.
For an internal action, this set consists of the automaton containing the internal action.
For an output action, this set consists of the automaton containing the output action and the automata that contain the input actions to which the output action is bound.
Two actions can be executed concurrently if their respective sets of state variables are disjoint.
The implied automata sets for each local action in Figure~\ref{sys_model} are: $out!(s) \to \{root, C\}$, $out!(t) \to \{C, D\}$, $process1 \to \{D\}$, $process2 \to \{D\}$, and $detect! \to \{D, root\}$.
Thus, $out!(s)$ and $process1$ can be executed concurrently while $process1$ and $detect!$ can not.

@@

\paragraph*{Parameters and action types}
Actions in the I/O automata model can be defined using parameters.
Parameters are useful for situations requiring fan-in or session semantics.
For example, consider an automaton with an input action $in?(s)$ that is to be bound to $N$ different output actions.
To get around the limitation that an input can only be bound to one output, we augment the input action with a parameter that indicates the associated output action, e.g., $in?[i](s)$ where $i$ takes on the integer values from 1 to $N$.
\emph{Parameterized} actions take a parameter while \emph{unparameterized} actions do not.
Parameters, for actions that require them, are specified in the binding and serve to identify the action.

@@

\paragraph*{Example}
Figure~\ref{example1} depicts a remote sensor system as a composition of automata using a scheme similar to Figure~\ref{sys_model} where parameters are given in brackets after the action name, e.g., \emph{recv?[0](m)}.
The Remote Sensor automaton consists of the Control automaton Z and three Channel automata A, B, and C representing three network connections.
The Control automaton Z consists of the Sensor automaton S, the Filter automaton F, and the Statistics automaton T.
The Control automaton C starts the sampling process with the uv-up output $Z.trigger!$.
The uv-up input $S.trigger?$ is executed atomically with $Z.trigger!$ due to the binding $(Z.trigger!, S.trigger?)$.
When the sampling process is complete, the v-up output $S.sample!(s)$ distributes the sample to the Filter F and Statistics component T via the $(S.sample!(s), F.in?(s))$ and $(S.sample!(s), T.in?(s))$ bindings.
Note that $S.sample!(s)$, $F.in?(s)$, and $T.in?(s)$ all agree on the type of the sample $s$.
The Filter automaton F passes the filtered sample to the Control automaton Z via the $(F.out!(t), Z.sample?(t))$ binding.
The Statistics automaton T calculates the statistics incrementally using the up internal $T.compute$.
The statistics can be polled by the Control automaton Z by issuing a request via the $(Z.request!, T.request?)$ binding and then receiving a response via the $(T.response!(u), Z.response?(u))$ binding.
The Control automaton Z contains a vp-output $Z.send![](m)$ and a vp-input $Z.recv?[](m)$ for sending and receiving network messages using the Channel automata A, B, and C.
The parameters 0, 1, and 2, are used to route messages to Channel automata A, B, and C respectively using the session idea mentioned previously.
Opportunities for concurrent execution exist for the remote sensor depicted in Figure~\ref{example1}.
The set of automata implied by $Z.trigger!$ is $\{Z, S\}$ while the set of automata implied by $T.compute$ is $\{T\}$.
Since the sets are disjoint, the two actions can be executed concurrently.
\fi

I/O automata thus provide a natural component abstraction and are therefore a suitable foundation for developing reusable components for asynchronous and concurrent software systems.
Each action is only allowed to modify the state of the component automaton with which it is associated, which serves to encapsulate behavior and avoid unexpected side effects.
The absence of shared state thus ensures that each component automaton can be deployed independently.
The external signature of a component automaton constitutes its interface and the result of interacting with it is well-defined since every action is atomic.
Composition, atomicity, asynchrony, and concurrency are thus well supported by the model itself.

@@

The I/O automata model assumes that systems are composed of a finite or countably infinite number of automata~\cite{lynch1996distributed}.
Assuming a countably infinite number of participants allows properties of the systems to be stated in terms of an abstract quantity representing the number of members, e.g., $N$, and is reasonable because all real systems are composed of a finite number of members.
The set of interactions in the I/O automata model is also fixed since automata are statically composed.
Dynamic systems can be \emph{modeled} in that approach by assuming that only a subset of the all the automata that could possibly exist are actively participating in the system.
To use this technique, a flag is associated with each automaton indicating if it is active or inactive and external actions for activation and deactivation are defined~\cite{lynch1994atomic}.

\paragraph*{Limitations of the static model}
While assuming a fixed number of automata is reasonable for modeling, assuming a fixed number of automata for a real system may result in either over-provisioning or inflexibility.
To develop a component based on the static I/O automata model, a developer must choose a concrete value for the abstract quantity $N$.
System resources are wasted if the actual number of participants is much smaller than $N$.
If the number of participants is greater than $N$, then the system cannot respond to certain situations even though resources might be available.
Thus, the techniques for \emph{modeling} a dynamic system of I/O automata are not necessarily appropriate for \emph{implementing} a dynamic system of I/O automata.
%% Consequently, we add the ability to deal with a dynamic set of automata and interactions.

To illustrate, let a system $S=(A,B)$ be a pair consisting of a set of automata $A$ and a set of bindings $B$.
The I/O automata model assumes that $S$ and therefore $A$ and $B$ are both fixed, so that dynamics must be \emph{simulated} via the manipulation of activation flags.

\paragraph*{Support for dynamic systems}
To address the limitations of the static representation of dynamic systems, ioa++ provides
architectural support for dynamic systems by separating the concerns of \emph{enacting} versus \emph{initiating} 
actions for managing dynamics into distinct \emph{system} and \emph{base} automata respectively.
We define four \emph{system actions} for managing dynamic systems of automata: \emph{create}, \emph{bind}, \emph{unbind}, and \emph{destroy}.
The \emph{create} action allows an automaton to create a child automaton.
The \emph{bind} action allows an automaton to associate an input action with an output action while the \emph{unbind} action allows an automaton to dissociate an input action from an output action.
The \emph{destroy} action allows an automaton to destroy a child automaton and all bindings associated with it (recursively).

@@

We note that system action requests may fail for a variety of reasons.
A create fails if the automaton that should be added to the set of automata, i.e., the automaton being created, already exists.
A bind fails if the automaton associated with the output action or the automaton associated with the input action does not exist, or if any of the binding rules outlined in Section~\ref{component_structure} would be violated.
An unbind fails if the binding does not exist.
Note that the owner of the binding is taken into account when determining whether the binding exists, such that only the owner of a binding can dissolve it.
A destroy fails if the automaton being destroyed does not exist or is not a child of the automaton requesting the destroy action.

@@

\paragraph*{Binding predicates}
The static configuration of a system as prescribed by the I/O automata model guarantees that the set of implied automata for an action is always the same.
This property is necessary and desirable for reasoning about the behavior of the system.
Admitting dynamics removes this guarantee since the set of implied automata will change according to the various system actions being executed.
To be specific, the set of implied automata for internal actions does not change but the set of implied automata for output actions can be changed with a successful bind, unbind, or destroy.

To allow a dynamic system to provide semantics consistent with a static system, the execution of output actions must be delayed until the set of implied automata matches the set implied by the static system.
To do this, we take the conjunction of the precondition of an output action with a predicate over the set of bindings $B$ (called a \emph{binding predicate}) to ensure that the output action is not executed until its set of implied automata is correct.
Thus, a dynamic system can be analyzed as a static system by ignoring system actions and the binding predicates.
A system defined statically can be implemented as a dynamic system by adding binding predicates and logic to create the required constellation of automata.

>>>>

The Input/Output (I/O) automata model was developed to model reactive concurrent and asynchronous systems~\cite{lynch1996distributed}.
An I/O automaton consists of state variables and a set of atomic actions that manipulate the state variables.
An action is only permitted to manipulate the state variables of the automaton to which it belongs.
An action is either an internal action, an output action, or an input action.
An internal action only changes the state of the automaton.
An output action changes the state of the automaton and produces a signal or value that may be consumed by one or more input actions.
An input action changes the state of the automaton when it receives a signal or value produced by an output action.
\ifjournal
The set of actions associated with an automaton is known as the automaton's \emph{signature}~\cite{lynch1996distributed}.
\fi
\ifjournal
Often, a local action consists of a precondition and an effect.
The precondition is a predicate over the state variables that enables the effect when selected for execution.
The effect is a function that computes the next state of the automaton from the current state.
An input action presented in this style consists solely of an effect.
\fi
I/O automata are \emph{composed} by (1)~concatenating the state variables of the constituent automata, (2)~consolidating the effects of all input actions with the same name, and (3)~incorporating the effects of input actions into output actions with the same name.
Execution in the I/O automata model consists of non-deterministically and repeatedly selecting a local action (an output or internal action~\cite{lynch1996distributed}) and then if the precondition is true applying the effect.
The scheduler is assumed to be \emph{fair,} meaning that a local action is guaranteed to be selected (but not necessarily executed) infinitely often.

@@

\begin{figure*}
\center
\includegraphics[width=\textwidth]{system_model}
\caption{Example event detection system.}
\label{sys_model}
\end{figure*}

\paragraph*{Motivation for I/O automata based components.}
%The I/O automata formalism was developed to model reactive systems~\cite{lynch1987hierarchical} and has been used in the design and verification of concurrent and distributed systems.
I/O automata are a good foundation for such components because they have independent state, straightforward interfaces, and precise semantics for event generation, distribution, and consumption resulting in well-defined interactions under composition.
\emph{Hierarchical decomposition}, the modeling dual of recursive encapsulation, is an established technique for I/O automata~\cite{lynch1994atomic} which allows a \emph{parent} to contain zero or more \emph{children}.
The \emph{root automaton} is the automaton at the top of the hierarchy and represents the entire system as composition is applied recursively up the hierarchy.

@@

\paragraph*{Explicit composition.}
Composition in the I/O automata model is defined at the automaton level and is based on actions having the same name.
To prepare for dynamic composition, we redefine composition at the action level.
An explicit association between an output action and input action is called a \emph{binding}.
%% The automaton that prescribes a binding is said to \emph{own} the binding.
%% A binding is denoted $(output, input, owner)$.
The set of bindings in a system must adhere to four essential rules.
First, the output action and input action of a binding must agree on the type being produced and consumed.
Second, an input action can be bound to at most one output action.
Third, an output action cannot be bound to more than one input action in the same automaton.
Fourth, an output action cannot be bound to an input action in the same automaton.
The second, third, and fourth rules exist because they admit ambiguity into the relationship between the states of the associated automata.
Graphically, an arrow originating in one component automaton must terminate in another component automaton as Figure~\ref{sys_model} illustrates.
%% To adhere to this rule, the following must be true for the global binding map $B$:  $\langle \forall o, i : (o, i) \in B :: \pi (o) \neq \pi (i) \rangle$.

\paragraph*{Concurrent execution.}
Recall that composing I/O automata consists of concatenating state variables and folding input actions into output actions and that the result of composition is a single equivalent automaton.
Since the state of each automaton is independent, opportunities exist for concurrent execution if the state of each automaton is preserved and a composition is not reduced to a single equivalent automaton.
Each local action implies a set of automata which in turn implies a set of state variables that might be modified.
For an internal action, this set consists of the automaton containing the internal action.
For an output action, this set consists of the automaton containing the output action and the automata that contain the input actions to which the output action is bound.
Two actions can be executed concurrently if their respective sets of state variables are disjoint.
The implied automata sets for each local action in Figure~\ref{sys_model} are: $out!(s) \to \{root, C\}$, $out!(t) \to \{C, D\}$, $process1 \to \{D\}$, $process2 \to \{D\}$, and $detect! \to \{D, root\}$.
Thus, $out!(s)$ and $process1$ can be executed concurrently while $process1$ and $detect!$ can not.

@@

\paragraph*{Parameters and action types.}
Actions in the I/O automata model can be defined using parameters.
Parameters are useful for situations requiring fan-in or session semantics.
For example, consider an automaton with an input action $in?(s)$ that is to be bound to $N$ different output actions.
To get around the limitation that an input can only be bound to one output, we augment the input action with a parameter that indicates the associated output action, e.g., $in?[i](s)$ where $i$ takes on the integer values from 1 to $N$.
\emph{Parameterized} actions take a parameter while \emph{unparameterized} actions do not.
Parameters, for actions that require them, are specified in the binding and serve to identify the action.

@@

\paragraph*{Example.}
Figure~\ref{example1} depicts a remote sensor system as a composition of automata using a scheme similar to Figure~\ref{sys_model} where parameters are given in brackets after the action name, e.g., \emph{recv?[0](m)}.
The Remote Sensor automaton consists of the Control automaton Z and three Channel automata A, B, and C representing three network connections.
The Control automaton Z consists of the Sensor automaton S, the Filter automaton F, and the Statistics automaton T.
The Control automaton C starts the sampling process with the uv-up output $Z.trigger!$.
The uv-up input $S.trigger?$ is executed atomically with $Z.trigger!$ due to the binding $(Z.trigger!, S.trigger?)$.
When the sampling process is complete, the v-up output $S.sample!(s)$ distributes the sample to the Filter F and Statistics component T via the $(S.sample!(s), F.in?(s))$ and $(S.sample!(s), T.in?(s))$ bindings.
Note that $S.sample!(s)$, $F.in?(s)$, and $T.in?(s)$ all agree on the type of the sample $s$.
The Filter automaton F passes the filtered sample to the Control automaton Z via the $(F.out!(t), Z.sample?(t))$ binding.
The Statistics automaton T calculates the statistics incrementally using the up internal $T.compute$.
The statistics can be polled by the Control automaton Z by issuing a request via the $(Z.request!, T.request?)$ binding and then receiving a response via the $(T.response!(u), Z.response?(u))$ binding.
The Control automaton Z contains a vp-output $Z.send![](m)$ and a vp-input $Z.recv?[](m)$ for sending and receiving network messages using the Channel automata A, B, and C.
The parameters 0, 1, and 2, are used to route messages to Channel automata A, B, and C respectively using the session idea mentioned previously.
Opportunities for concurrent execution exist for the remote sensor depicted in Figure~\ref{example1}.
The set of automata implied by $Z.trigger!$ is $\{Z, S\}$ while the set of automata implied by $T.compute$ is $\{T\}$.
Since the sets are disjoint, the two actions can be executed concurrently.
\fi

@@

%% The I/O automata model assumes that systems are composed of a finite or countably infinite number of automata~\cite{lynch1996distributed}.
%% Assuming a countably infinite number of participants allows properties of the systems to be stated in terms of an abstract quantity representing the number of members, e.g., $N$, and is reasonable because all real systems are composed of a finite number of members.
%% The set of interactions in the I/O automata model is also fixed since automata are statically composed.
%% Dynamic systems can be \emph{modeled} in that approach by assuming that only a subset of the all the automata that could possibly exist are actively participating in the system.
%% To use this technique, a flag is associated with each automaton indicating if it is active or inactive and external actions for activation and deactivation are defined~\cite{lynch1994atomic}.

%% \paragraph*{Limitations of the static model.}
%% While assuming a fixed number of automata is reasonable for modeling, assuming a fixed number of automata for a real system may result in either over-provisioning or inflexibility.
%% To develop a component based on the static I/O automata model, a developer must choose a concrete value for the abstract quantity $N$.
%% System resources are wasted if the actual number of participants is much smaller than $N$.
%% If the number of participants is greater than $N$, then the system cannot respond to certain situations even though resources might be available.
%% Thus, the techniques for \emph{modeling} a dynamic system of I/O automata are not necessarily appropriate for \emph{implementing} a dynamic system of I/O automata.
%% %% Consequently, we add the ability to deal with a dynamic set of automata and interactions.

%% To illustrate, let a system $S=(A,B)$ be a pair consisting of a set of automata $A$ and a set of bindings $B$.
%% The I/O automata model assumes that $S$ and therefore $A$ and $B$ are both fixed, so that dynamics must be \emph{simulated} via the manipulation of activation flags.

\paragraph*{Support for dynamic systems.}
The I/O automata model assumes that systems are composed of a finite or countably infinite number of automata~\cite{lynch1996distributed} and models dynamic systems by activating and deactivating automata as appropriate~\cite{lynch1994atomic}.
While appropriate for modeling, this assumption is not conducive to building flexible and efficient systems because it requires \emph{a priori} bounds on all resources.
To address the limitations of the static representation of dynamic systems, ioa++ provides architectural support for dynamic systems by separating the concerns of \emph{enacting} versus \emph{initiating} actions for managing dynamics into distinct \emph{system} and \emph{base} automata respectively.
We define four \emph{system actions} for managing dynamic systems of automata: \emph{create}, \emph{bind}, \emph{unbind}, and \emph{destroy}.
The \emph{create} action allows an automaton to create a child automaton.
The \emph{bind} action allows an automaton to associate an input action with an output action while the \emph{unbind} action allows an automaton to dissociate an input action from an output action.
The \emph{destroy} action allows an automaton to destroy a child automaton and all bindings associated with it (recursively).

@@

@@

\paragraph*{Binding predicates.}
Consider a simple system consisting of two automata where automaton A reliably transfers messages to automaton B.
In a dynamic system, one cannot guarantee that the message will be received because there is no way to determine if the output action producing the message is bound to the input action receiving the message.
Since we can guarantee that the set of binding does not change during the execution of an action, we allow automata to query the current set of bindings using a \emph{binding predicate}.
Binding predicates are typically used the precondition of output actions to guarantee delivery.

### 2012-02-23 papers/paper2/conclusion.tex
Removing material.

I/O automata offer a good foundation for asynchronous and concurrent components due to their support for independent state, well-defined interfaces, and well-defined interactions under composition.
The ioa++ framework facilitates concurrent execution via dynamic composition and the degree of concurrency is limited only by the interactions of the automata and the overhead of the framework.
The two key challenges when moving from the formal model to an actual implementation were introducing features for managing dynamic constellations of automata and providing a concrete scheduling mechanism.
We agree with~\cite{georgiou2009automated} that having a representation of a model that can be compiled is beneficial because it allows one to reason about the program directly from the source code and helps to close the gap between the developer's mental model and the code listing.
Our primary goal moving forward is to use the ioa++ framework to gain further experience building real systems with I/O automata, and to support optimization of its performance with respect to an increasingly diverse range of use cases.

>>>>

I/O automata are a good basis for asynchronous and concurrent components due to their support for independent state, well-defined interfaces, and well-defined interactions under composition.
The ioa++ framework facilitates concurrent execution via dynamic composition and the degree of concurrency is limited only by the interactions of the automata and the overhead of the framework.
Two key challenges when moving from the formal model to an actual implementation were introducing features for managing dynamic constellations of automata and providing a concrete scheduling mechanism.
We agree with~\cite{georgiou2009automated} that having a model that can be compiled is beneficial for reasoning about the program directly from the source code and helps to close the gap between the developer's mental model and the code listing.
Our primary goal moving forward is to use the ioa++ framework to gain further experience building real systems with I/O automata, and to support optimization of its performance with respect to an increasingly diverse range of use cases.
